---
title: "Hyperosmotic Stress in Eukaryotic Cells"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Hyperosmotic Stress in Eukaryotic Cells}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: cpdat.bib
csl: peerj.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available (R-Forge?)
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette from the R package [canprot](https://github.com/jedick/canprot) version `r packageDescription("canprot")$Version` shows compositional metrics and phylostrata for proteins that are differentially expressed in hyperosmotic compared to control conditions.
Abbreviations:

  * <i>Z</i><sub>C</sub> &ndash; carbon oxidation state; <i>n</i><sub>H<sub>2</sub>O</sub> &ndash; stoichiometric hydration state; <i>n</i><sub>O<sub>2</sub></sub> &ndash; stoichiometric oxidation state.
  * <i>n</i><sub>AA</sub> &ndash; protein length; PS &ndash; phylostrata.
  * <i>n</i><sub>down</sub> &ndash; number of down-regulated proteins; <i>n</i><sub>up</sub> &ndash; number of up-regulated proteins.

Stoichiometric values are calculated using basis species (rQEC derivation) or amino acid biosynthetic reactions ([<span title="Distinct trends in chemical composition of proteins from metagenomes in redox and salinity gradients">Dick et al., 2020</span>](https://doi.org/10.1101/2020.04.01.020008); [<span title="Water as a reactant in the differential expression of proteins in cancer">Dick, 2020</a>](https://doi.org/10.1101/2020.04.09.035022)).
Gene ages, expressed as phylostrata numbers, are taken from [<span title="Altered interactions between unicellular and multicellular genes drive hallmarks of transformation in a diverse range of solid tumors">Trigos et al., 2017</span>](https://doi.org/10.1073/pnas.1617743114) (TPPG17) or [<span title="Towards consensus gene ages ">Liebeskind et al, 2016</span>](https://doi.org/10.1093/gbe/evw113) (LMM16).

```{r options, echo=FALSE}
options(width = 90)
```

<style type="text/css">
body {
  max-width: 800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r datasets}
datasets <- pdat_osmotic_euk(2020)
```

```{r comptab, results="hide", message=FALSE, echo = FALSE}
pdat1 <- lapply(datasets, pdat_osmotic_euk)
comptab1 <- lapply(pdat1, get_comptab)
pdat2 <- lapply(pdat1, recomp, "biosynth")
comptab2 <- lapply(pdat2, get_comptab, "nO2")
comptab3 <- lapply(pdat1, get_comptab, "nAA", "PS")
comptab4 <- lapply(pdat1, get_comptab, "nAA", "PS", PS_source = "LMM16")
```

Dashed contour lines in the plots outline the 50% credible region for highest probability density.

```{r diffplot, fig.width=8, fig.height=8, fig.align = "center", echo = FALSE, pngquant = pngquant}
par(mfrow = c(2, 2), mar = c(4, 4, 2, 2), mgp = c(2.5, 1, 0))
diffplot(comptab1)
title(quote("rQEC"~italic(n)[H[2]*O]), font.main = 1)
diffplot(comptab2, c("nO2", "nH2O"))
title(quote("Biosynthetic"~italic(n)[H[2]*O]~"and"~italic(n)[O[2]]), font.main = 1)
diffplot(comptab3, c("nAA", "PS"))
title("Trigos et al. (2017) ages", font.main = 1)
diffplot(comptab4, c("nAA", "PS"))
title("Liebeskind et al. (2016) ages", font.main = 1)
```

In the table, values of &Delta;<i>Z</i><sub>C</sub>, &Delta;<i>n</i><sub>H<sub>2</sub>O</sub> and &Delta;<i>n</i><sub>O<sub>2</sub></sub> are multiplied by 1000, values of &Delta;PS are multiplied by 100, and negative values are shown in bold.

```{r xsummary, results="asis", echo = FALSE}
library(xtable)
out <- xsummary2(comptab1, comptab2, comptab3, comptab4)
# round values and include dataset tags
tags <- sapply(sapply(strsplit(datasets, "="), "[", -1), paste, collapse = ";")
out <- cbind(out[, 1:2], tags = tags, out[, 3:25])
out[, 6:26] <- round(out[, 6:26], 4)
write.csv(out, "osmotic.csv", row.names = FALSE, quote = 2)
```

## Data Sources

__a__. Dataset S3 of @LTH+11, filtered to include proteins that are quantified in at least 14 of 15 experiments (5 time points with 3 replicates each) and have the same direction of change in all experiments.
__b__. Supplementary Table 1 of @OBBH11.
__c__. __d__. __e__. __f__. __g__. __h__. Supplementary Table S1 of @LFY+12 (sheet "All proteins"), filtered to include proteins with _q_-value < 0.1.
__i__. Table 2 of @CLG+15.
__j__. Table 1 of @YDZ+15.
__k__. Supplementary Tables 1 and 2 of @JBG+18.
__l__. __m__. Supplementary Table S1 of @SMS+18.

## References
