---
title: "Hyperosmotic Stress in Bacteria"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Hyperosmotic Stress in Bacteria}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: cpdat.bib
csl: peerj.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available (R-Forge?)
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette from the R package [canprot](https://github.com/jedick/canprot) version `r packageDescription("canprot")$Version` shows compositional metrics and phylostrata for proteins that are differentially expressed in hyperosmotic compared to control conditions.
Abbreviations:

  * <i>Z</i><sub>C</sub> &ndash; carbon oxidation state; <i>n</i><sub>H<sub>2</sub>O</sub> &ndash; stoichiometric hydration state; <i>n</i><sub>O<sub>2</sub></sub> &ndash; stoichiometric oxidation state.
  * <i>n</i><sub>AA</sub> &ndash; protein length.
  * <i>n</i><sub>down</sub> &ndash; number of down-regulated proteins; <i>n</i><sub>up</sub> &ndash; number of up-regulated proteins.

Stoichiometric values are calculated using basis species (rQEC derivation) or amino acid biosynthetic reactions ([<span title="Distinct trends in chemical composition of proteins from metagenomes in redox and salinity gradients">Dick et al., 2020</span>](https://doi.org/10.1101/2020.04.01.020008); [<span title="Water as a reactant in the differential expression of proteins in cancer">Dick, 2020</a>](https://doi.org/10.1101/2020.04.09.035022)).

```{r options, echo=FALSE}
options(width = 90)
```

<style type="text/css">
body {
  max-width: 800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r datasets}
datasets <- pdat_osmotic_bact(2020)
```

```{r comptab, results="hide", message=FALSE, echo = FALSE}
pdat1 <- lapply(datasets, pdat_osmotic_bact)
comptab1 <- lapply(pdat1, get_comptab)
pdat2 <- lapply(pdat1, recomp, "biosynth")
comptab2 <- lapply(pdat2, get_comptab, "nO2")
comptab3 <- lapply(pdat1, get_comptab, "nAA", "GRAVY")
comptab4 <- lapply(pdat1, get_comptab, "pI", "GRAVY")
```

Dashed contour lines in the plots outline the 50% credible region for highest probability density.

```{r diffplot, fig.width=8, fig.height=4, fig.align = "center", echo = FALSE, pngquant = pngquant}
par(mfrow = c(1, 2), mar = c(4, 4, 2, 2), mgp = c(2.5, 1, 0))
diffplot(comptab1)
title(quote("rQEC"~italic(n)[H[2]*O]), font.main = 1)
diffplot(comptab4, c("pI", "GRAVY"))
```

In the table, values of &Delta;<i>Z</i><sub>C</sub>, &Delta;<i>n</i><sub>H<sub>2</sub>O</sub>, &Delta;<i>n</i><sub>O<sub>2</sub></sub> and &Delta;GRAVY are multiplied by 1000, values of &Delta;pI are multiplied by 100, and negative values are shown in bold.

```{r xsummary, results="asis", echo = FALSE}
library(xtable)
out <- xsummary3(comptab1, comptab2, comptab3, comptab4)
# round values and include dataset tags
tags <- sapply(sapply(strsplit(datasets, "="), "[", -1), paste, collapse = ";")
out <- cbind(out[, 1:2], tags = tags, out[, 3:25])
out[, 6:26] <- round(out[, 6:26], 4)
write.csv(out, "osmotic_bact.csv", row.names = FALSE, quote = 2)
```

## Data Sources

__a__. Additional file 3: Table S2 of @PNWB09.
__b__. Supplementary Table 8 of @FTR+10. Only proteins with consistent expression ratios (all > 1 or all < 1) at each time point (15, 60, and 180 min.) were included.
__c__. Supporting Information Table 1 of @LPK+13 (sheets "Up-Down Proteins" and "Unknown function").
__d__. __e__. Supplementary Tables S3A and S3B of @QHT+13.
__f__. Table 1 (proteins) and supplemental Table S2 (genes) of @ADW+14.
__g__. __h__. __i__. __j__. Supporting Information Table S2 of @KKG+14.
__k__. __l__. Supporting Information Table of @PBP+14, filtered to include proteins with _p_-value < 0.05.
__m__. __n__. Additional file Table S2 of @KLB+15.
__o__. __p__. Supplementary Table S6 of @SKV+16, filtered to include proteins with fold change > 2 or < 0.5 for the ratios Glucose / LB (lysogeny broth) or Osmotic-stress glucose / LB.
__q__. Supplementary Table 1 (sheets "0.76 fold down regulated" and "1.3 fold up regulated") of @KAK+17.
__r__. Supplemental Table S-2 of @LYS+17, filtered to include proteins with log~2~ fold change > 1 or < -1 and _p_-value < 0.05.
__s__. Supplementary Table 1 of @KSK+18 (amino acid compositions computed from protein sequences in the list of gene annotations).
__t__. __u__. Tables S1–S6 of @LJC+18. For each of the wild-type and ∆_sigB_ mutant, only proteins that were identified in multicellular vesicles in a single condition (0.5 M salt stress or without salt stress) were included.
__v__. __w__. Extracted from proteinGroups.txt in PRIDE project [PXD010072](https://www.ebi.ac.uk/pride/archive/projects/PXD010072)/MaxQuantOutput.tar.gz [@TSC18], filtered to include proteins with non-zero LFQ intensity values for all replicates in each experiment; the medians of these values were used to compute fold changes; proteins with fold change > 1.5 or < 2/3 were kept.
__x__. Table 2 of @LWS+19.
__y__. __z__. Supplementary Tables S4 and S5 of @MGF+19, filtered to include proteins with fold change >= 2 or <= 0.5.
__A__. Supplementary Table 1 (sheets ">2.0 Fold" and "< 0.5 Fold") of @AST+20.
__B__. __C__. Supplementary Table 1 of @GBR+20 (column "MMO+NaCl/MMO" for CIRM129 and CIRM1025).

## References
