---
title: "3D Cell Culture"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{3D Cell Culture}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: cpdat.bib
csl: peerj.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available (R-Forge?)
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette shows compositional metrics and phylostrata for proteins that are differentially expressed in 3D compared to 2D culture conditions.
Abbreviations:

  * <i>Z</i><sub>C</sub> &ndash; carbon oxidation state; <i>n</i><sub>H<sub>2</sub>O</sub> &ndash; stoichiometric hydration state; <i>n</i><sub>O<sub>2</sub></sub> &ndash; stoichiometric oxidation state.
  * <i>n</i><sub>AA</sub> &ndash; protein length; PS &ndash; phylostrata.
  * <i>n</i><sub>down</sub> &ndash; number of down-expressed proteins; <i>n</i><sub>up</sub> &ndash; number of up-expressed proteins.

Stoichiometric values are derived from basis species using the rQEC method (Dick et al., 2020) or from amino acid biosynthetic reactions.
References for gene ages: [Trigos et al. (2017)](https://doi.org/10.1073/pnas.1617743114) (TPPG17); [Liebeskind et al. (2016)](https://doi.org/10.1093/gbe/evw113) (LMM16).

```{r options, echo=FALSE}
options(width = 90)
```

<style type="text/css">
body {
  max-width: 800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r datasets}
datasets <- pdat_3D()
```

```{r comptab, results="hide", message=FALSE, echo = FALSE}
pdat1 <- lapply_canprot(datasets, pdat_3D)
comptab1 <- lapply(pdat1, get_comptab)
pdat2 <- lapply(pdat1, pdat_recomp, "biosynth")
comptab2 <- lapply(pdat2, get_comptab, "nO2")
comptab3 <- lapply(pdat1, get_comptab, "nAA", "PS")
comptab4 <- lapply(pdat1, get_comptab, "nAA", "PS", PS_source = "LMM16")
```

Dashed contour lines in the plots outline the 50% credible region for highest probability density.
Triangle symbols indicate transcriptomic datasets; all others are protein expression.

```{r diffplot, fig.width=8, fig.height=8, fig.align = "center", echo = FALSE, pngquant = pngquant}
par(mfrow = c(2, 2), mar = c(4, 4, 2, 2), mgp = c(2.5, 1, 0))
pch <- ifelse(grepl("transcript", datasets), 2, 1)
diffplot(comptab1, pch = pch)
title(quote("rQEC"~italic(n)[H[2]*O]), font.main = 1)
diffplot(comptab2, c("nO2", "nH2O"), pch = pch)
title(quote("Biosynthetic"~italic(n)[H[2]*O]~"and"~italic(n)[O[2]]), font.main = 1)
diffplot(comptab3, c("nAA", "PS"), pch = pch)
title("Trigos et al. (2017) ages", font.main = 1)
diffplot(comptab4, c("nAA", "PS"), pch = pch)
title("Liebeskind et al. (2016) ages", font.main = 1)
```

In the table, values of &Delta;<i>Z</i><sub>C</sub>, &Delta;<i>n</i><sub>H<sub>2</sub>O</sub> and &Delta;<i>n</i><sub>O<sub>2</sub></sub> are multiplied by 1000, values of &Delta;PS are multiplied by 100, and negative values are shown in bold.

```{r xsummary, results="asis", echo = FALSE}
library(xtable)
out <- xsummary2(comptab1, comptab2, comptab3, comptab4)
# round values and include dataset tags
tags <- sapply(sapply(strsplit(datasets, "="), "[", -1), paste, collapse = ";")
out <- cbind(out[, 1:2], tags = tags, out[, 3:25])
out[, 6:26] <- round(out[, 6:26], 4)
write.csv(out, "3D.csv", row.names = FALSE, quote = 2)
```

## Data Sources
Gene names or other identifiers were converted to UniProt accession numbers using the <a href="https://www.uniprot.org/mapping/">UniProt mapping tool</a>.

__a__. Gene names from Supporting Information Table 1S of @PLC+10.
__b__. __c__. Sheets 2 and 3 in Table S1 of @MHG+12 (P2: passage 2; P5: passage 5).
__d__. __e__. Supplemental Table 1C of @MVC+12, filtered to include proteins with expression ratios < 0.77 or > 1.3.
__f__. Supplemental Table S5 of @YYW+13.
__g__. __h__. Supplemental Table S4 of @ZMH+13.
__i__. Table S1 of @HKX+14.
__j__. Table S2 of @RKP+14, filtered to include proteins that have differences in spectral counts recorded in at least two of three experiments, absolute overall fold change is ≥ 1.5 or ≤ 2/3, and _p_-value is < 0.05.
__k__. P1&lowbar;Data sheet in the Supporting Information file of @WRK+14.
__l__. Supplemental Table S1 of @MTK+15, filtered to exclude marked contaminants and reverse sequences and to include proteins with "Ratio H/L normalized" > 1.5 or < 2/3.
__m__. Supplemental Dataset 2 of @MDN+16, filtered to include genes with log~2~ fold change > 1 or < -1 and _p_-value < 0.05.
__n__. Gene names from Table II of @SSPR16.
__o__. Tables S1a and S1b of @YLW+16.
__p__. Supporting Information S2 Table of @PPM+17, filtered to include transcripts with "fold change (3D treatment, 2D control)" > 2 or < -2 and "Adjusted p-Value for Diff of Treatment = (2D) - (3D)" < 0.05.
__q__. Tables S2 and S3 of @KJK+18.
__r__. __s__. Supplementary Table 3 of @TGD18. 
__t__. Extracted from Supplementary Table 1 of @GADS19. Values were quantile normalized, then ratios were calculated between 3D and 2D cultures for each treatment (control, CBD, UVA, UVA+CBD, UVB, UVB+CBD). Ratios > 1.2 or < 1/1.2 in at least 4 treatments were used to identify differentially expressed proteins.
__u__. Supplementary Data I, sheet "Volcano plot 2Dv3D" of @HLC19.
__v__. __w__. __x__. Supporting Table 1 of @LPK+19, sheets "3P2P" (mono-cultured preadipocytes), "3A2A" (mono-cultured adipocytes), "3C2C" (co-cultured adipocytes with macrophages), filtered to include proteins marked as "T-test Significant" and with absolute value of "N: T-test Difference" > log~2~(0.5).

## References
