---
title: "canprot vignette"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{canprot vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(canprot)
options(width = 72)
```

This is a generic vignette for the canprot package.
To save package space and checking time, prebuilt vignettes are not included in the package.

To compile the vignettes on demand and view them in the browser, use the `mkvig()` function.
For example, `mkvig("3D")` compiles the vignette for three-dimensional cell culture and then opens it in the browser.
Each of the vignettes is also available as a demo, which can be run from the browser (`help.start()` > Packages > canprot > Code demos).

Here is a list of available vignettes:
```{r vignettes, echo = FALSE}
functions <- grep("pdat_", ls("package:canprot"), value = TRUE)
vignettes <- gsub("pdat_", "", functions)
vignettes
```

Note that building the vignettes requires pandoc as a system dependency.
See the [rmarkdown pandoc vignette](https://CRAN.R-project.org/package=rmarkdown/vignettes/pandoc.html) for more information.

The vignettes can be viewed online at <http://chnosz.net/canprot/doc/index.html>.
