---
title: "Demo for canprot"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Demo for canprot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: canprot.bib
link-citations: yes
csl: elementa.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
oldopt <- options(width = 80)
```

```{r HTML, include=FALSE}
Zc <- "<i>Z</i><sub>C</sub>"
```

The **canprot** package calculates chemical metrics of proteins from amino acid compositions.
This vignette was compiled on `r Sys.Date()` with **canprot** version `r packageDescription("canprot")$Version`.

Next vignette: [Introduction to canprot](introduction.html)

## canprot demo

Run the demo using `demo("canprot")`.
The code and output of the demo are shown below.

The **canprot** functions used are:

- `human.aa()`: Gets amino acid compositions of human proteins from UniProt IDs
- `plength()`: Calculates protein length (this line is commented out)
- `Zc()`: Calculates carbon oxidation state
- `pI()`: Calculates isoelectric point
- `add.cld()`: Adds compact letter display to a boxplot

```{r echo = FALSE}
knitr::read_chunk("../demo/canprot.R")
```

```{r canprot_demo_setup, echo = FALSE}
```

```{r canprot_demo_body, out.width = "100%", fig.align = "center", fig.width = 8, fig.height = 4.5}
```

The plots show carbon oxidation state (`r Zc`) and isoelectric point (pI) for human proteins in different subcellular locations.
The localization data is from Table S6 of @TAW+17, filtered to include proteins that have both a validated location and only one predicted location.

```{r reset, include=FALSE}
options(oldopt)
```

## References
