---
title: "Hypoxia and 3D Culture"
output: html_vignette
vignette: >
  %\VignetteIndexEntry{Hypoxia and 3D Culture}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: data_sources.bib
csl: peerj.csl
---

## Data Sources

[`HXS+06`](http://doi.org/10.1002/pmic.200500754) 2% O<sub>2</sub> vs normoxic conditions. Source: Table 1 of @HXS+06.

[`BRA+10`](http://doi.org/10.1177/1933719109351320) 1% vs 6% O<sub>2</sub>. Source: Tables 2 and 3 of @BRA+10.

[`DPL+10`](http://doi.org/10.1021/pr900829h) The comparisons here use expression ratios HYP/LSC (oxygen deprivation / low serum control) >1.2 or <0.83, calculated from the reported ratios. Source: extracted from Suppl. Table 2 of @DPL+10, including proteins with _p_-value <0.05 and EF <1.4.

[`BMJ+11`](http://doi.org/10.1016/j.radonc.2011.05.058) Translationally regulated genes. Source: Suppl. Tables 1–4 of @BMJ+11.

[`CBW+11`](http://doi.org/10.1021/pr100982j) 1% O<sub>2</sub> for 72 h vs standard conditions. Source: Suppl. Table 1(a) of @CBW+11.

[`LAR+12`](http://doi.org/10.1021/pr201025m) Hypoxic vs control conditions for 16 h. Source: Suppl. Table S5 of @LAR+12.

[`MHG+12`](http://doi.org/10.1371/journal.pone.0052692) Tumourspheres (50 to 200 μm diameter) at passage 5 (P5) or 2 (P2) compared to adherent cells. Source: Sheets 2 and 3 in Table S1 of @MHG+12.

[`MVC+12`](http://doi.org/10.1021/pr2012472) Perinecrotic and necrotic regions compared to surface of multicell spheroids (~600 μm diameter). The comparisons here use expression ratios <0.77 or >1.3. Source: Suppl. Table 1C of @MVC+12.

[`FWH+13`](http://doi.org/10.1016/j.bbapap.2013.09.023) Incubation for several days under hypoxia (1% O<sub>2</sub>). Source: Suppl. Table 2A of @FWH+13 (control virus cells).

[`RHD+13`](http://doi.org/10.1074/mcp.M112.018325) Source: extracted from Suppl. Table 1 of @RHD+13, including proteins with iTRAQ ratios <0.83 or >1.2 and _p_-value <0.05.

[`VTMF13`](http://doi.org/10.1021/pr400758d) 5% O<sub>2</sub> vs atmospheric levels of O<sub>2</sub>. The comparisons here include proteins with a normalized expression ratio of >1.2 or <0.83. Source: SI table of @VTMF13.

[`DYL+14`](http://doi.org/10.1074/mcp.M114.038232) The comparisons here include proteins with _p_ <0.05. Source: Suppl. Table S1 of @DYL+14.

[`RKP+14`](http://doi.org/10.1186/1477-5956-12-39) Organotypic spheroids (~250 μm diameter) vs lysed CRC tissue. Source: extracted from Table S2 of @RKP+14, filtered as follows: at least two of three experiments have differences in spectral counts, absolute overall fold change is at least 1.5, and _p_-value is less than 0.05.

[`WRK+14`](http://doi.org/10.1371/journal.pone.0106973) SPH vs classical cell culture (2D growth). Standard concentrations of gases used for tissue culture (5% CO<sub>2</sub>, 95% air) were used in both cases. The comparisons here include proteins that have a log<sub>2</sub> fold change of at least ±1. Source: P1&lowbar;Data sheet in the SI of @WRK+14.

[`BSA+15`](http://doi.org/10.1155/2015/678527) 1% vs 19% O<sub>2</sub>. Source: Table S1 of @BSA+15.

[`HWA+16`](http://doi.org/10.1016/j.celrep.2016.01.036) 1% O<sub>2</sub> for 24 hr. The comparisons here include proteins with a fold change of <0.5 or >1 and proteins that were detected in only hypoxic or only normoxic conditions. Source: Table S1 of @HWA+16.

[`LCS16`](http://doi.org/10.1371/journal.pone.0153627) Microarray analysis of differential gene expression in the transcriptome (total rRNA) and translatome (polysomal / total RNA ratio) of cells grown in normal and hypoxic (1% O<sub>2</sub>) conditions. Source: data file supplied by Dr. Ming-Chih Lai (@LCS16).

[`RSE+16`](http://doi.org/10.1186/s13287-016-0310-7) ASC from 3 donors cultured for 24 hr. in hypoxic (1% O<sub>2</sub>) vs normoxic (20% O<sub>2</sub>) conditions. Source: Tables 1 and 2 of @RSE+16.

[`XCJ+16`](http://doi.org/10.1016/j.jprot.2015.09.028) Rat cardiomyocytes treated with CoCl<sub>2</sub> (hypoxia mimetic) vs control or with SAL (anti-hypoxic) vs CoCl<sub>2</sub>. Source: SI Tables 1S and 2S of @XCJ+16.

[`YLW+16`](http://doi.org/10.1021/acs.jproteome.6b00342) 800 μm spheroids vs 2D monolayers. Source: Tables S1a–b of @YLW+16.

## Summary Table

This table compares the chemical compositions of groups of proteins that are relatively down- and up-expressed (`n1` and `n2`, respectively) in cells grown in hypoxia or 3D culture compared to control conditions.

```{r options, echo=FALSE}
options(width = 90)
```

```{r canprot, message=FALSE}
library(canprot)
data(canprot)
```

```{r datasets}
datasets <- pdat_hypoxia()
```

```{r comptab, results="hide", message=FALSE}
comptab <- lapply_canprot(datasets, function(dataset) {
  pdat <- get_pdat(dataset, "pdat_hypoxia")
  ZC_nH2O(pdat, plot.it=FALSE)
}, varlist="pdat_hypoxia")
```

```{r xsummary, results="asis"}
library(xtable)
xsummary(comptab)
```

## Mean Differences

The reoxygenation or anti-hypoxic, tumor spheroid, and adipose-derived stem cell datasets are highlighted in blue, red, and orange, respectively.

```{r diffplot, fig.width=6, fig.height=6, fig.align="center"}
col <- rep("black", length(datasets))
col[grepl("ReOx", datasets)] <- "blue"
col[grepl("=SPH", datasets)] <- "red"
col[grepl("=ASC", datasets)] <- "orange"
diffplot(comptab, col=col)
```

## References
