---
title: "Lung Cancer"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Lung Cancer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: cpdat.bib
csl: peerj.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available (R-Forge?)
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette shows compositional metrics and phylostrata for proteins that are differentially expressed in lung cancer compared to normal tissue.
Abbreviations:

  * <i>Z</i><sub>C</sub> &ndash; carbon oxidation state; <i>n</i><sub>H<sub>2</sub>O</sub> &ndash; stoichiometric hydration state; <i>n</i><sub>O<sub>2</sub></sub> &ndash; stoichiometric oxidation state.
  * <i>n</i><sub>AA</sub> &ndash; protein length; PS &ndash; phylostrata.

Stoichiometric values are derived from basis species using the rQEC method (Dick et al., 2020) or from amino acid biosynthetic reactions.
References for gene ages: [Trigos et al. (2017)](https://doi.org/10.1073/pnas.1617743114) (TPPG17); [Liebeskind et al. (2016)](https://doi.org/10.1093/gbe/evw113) (LMM16).

```{r options, echo=FALSE}
options(width = 90)
```

<style type="text/css">
body {
  max-width: 800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r datasets}
datasets <- pdat_lung()
```

```{r comptab, results="hide", message=FALSE, echo = FALSE}
pdat1 <- lapply_canprot(datasets, pdat_lung)
comptab1 <- lapply(pdat1, get_comptab)
pdat2 <- lapply(pdat1, pdat_recomp, "biosynth")
comptab2 <- lapply(pdat2, get_comptab, "nO2")
comptab3 <- lapply(pdat1, get_comptab, "nAA", "PS")
comptab4 <- lapply(pdat1, get_comptab, "nAA", "PS", PS_source = "LMM16")
```

Dashed contour lines in the plots outline the 50% credible region for highest probability density.
Triangle symbols indicate transcriptomic datasets; all others are protein expression.

```{r diffplot, fig.width=8, fig.height=8, fig.align = "center", echo = FALSE, pngquant = pngquant}
par(mfrow = c(2, 2), mar = c(4, 4, 2, 2), mgp = c(2.5, 1, 0))
pch <- ifelse(grepl("transcript", datasets), 2, 1)
diffplot(comptab1, pch = pch)
title(quote("rQEC"~italic(n)[H[2]*O]), font.main = 1)
diffplot(comptab2, c("nO2", "nH2O"), pch = pch)
title(quote("Biosynthetic"~italic(n)[H[2]*O]~"and"~italic(n)[O[2]]), font.main = 1)
diffplot(comptab3, c("nAA", "PS"), pch = pch)
title("Trigos et al. (2017) ages", font.main = 1)
diffplot(comptab4, c("nAA", "PS"), pch = pch)
title("Liebeskind et al. (2016) ages", font.main = 1)
```

In the table, values of &Delta;<i>Z</i><sub>C</sub>, &Delta;<i>n</i><sub>H<sub>2</sub>O</sub> and &Delta;<i>n</i><sub>O<sub>2</sub></sub> are multiplied by 1000, values of &Delta;PS are multiplied by 100, and negative values are shown in bold.
Abbreviations:

  * <i>n</i><sub>down</sub> &ndash; number of down-expressed proteins; <i>n</i><sub>up</sub> &ndash; number of up-expressed proteins.
  * NSCLC &ndash; non-small cell lung cancer; SCC &ndash; squamous cell carcinoma; ADC &ndash; adenocarcinoma.
  * ANT &ndash; adjacent or patient-matched normal (non-malignant, benign) tissue; NBE &ndash; normal bronchial epithelium.
  * LCM &ndash; laser capture microdissection; FFPE &ndash; formalin-fixed paraffin-embedded.

```{r xsummary, results="asis", echo = FALSE}
library(xtable)
out <- xsummary2(comptab1, comptab2, comptab3, comptab4)
# round values and include dataset tags
tags <- sapply(sapply(strsplit(datasets, "="), "[", -1), paste, collapse = ";")
out <- cbind(out[, 1:2], tags = tags, out[, 3:11])
out[, 6:12] <- round(out[, 6:12], 4)
write.csv(out, "lung.csv", row.names = FALSE, quote = 2)
```

## Data Sources
Where given, gene names or other identifiers were converted to UniProt accession numbers using the <a href="https://www.uniprot.org/mapping/">UniProt mapping tool</a>, except for IPI accession numbers, which were converted using the <a href="https://david-d.ncifcrf.gov/conversion.jsp">DAVID 6.7 conversion tool</a>.

<b>a</b>. Table 1 of @LXC+06.
<b>b</b>. <b>c</b>. Gene names from Tables II and III of @KHA+12.
<b>d</b>. Gene names from Table 2 of @YLL+12.
<b>e</b>. IPI numbers from Table 1 of @ZZD+12.
<b>f</b>. Table 2 of @ZZY+13.
<b>g</b>. Table 1 of @LLY+14.
<b>h</b>. Supplementary data 1b of @LWT+14.
<b>i</b>. Supplementary Data of @ZLH+14.
<b>j</b>. Gene names from Supporting Information Table S5 of @ZLS+14 (differential expression between normal endothelial cells and both paratumor and tumor endothelial cells).
<b>k</b>. Table S3 of @KNT+15 (LPIA: lepidic predominant invasive adenocarcinoma vs pN: pseudo-normal), filtered to include proteins with log<sub>2</sub> fold change > 1 or < -1 and _p_-value < 0.05.
<b>l</b>. <b>m</b>. Gene names from Table S1 (genes) and UniProt names from Table S2 (proteins) of @BLL+16, filtered to include genes or proteins with log<sub>2</sub> fold change > 1.5 or < -1/1.5 and _p_-value < 0.05.
<b>n</b>. Gene names from Table S2 of @FGP+16.
<b>o</b>. Gene names from Table 1 of @JCP+16.
<b>p</b>. <b>q</b>. <b>r</b>. Supplemental Table S3B of @HHH+16 (primary tumor / normal comparison; pN0: no nodes involved; pN1: ipsilateral peribronchial/interlobar/hilar LN metastasis; pN2: ipsilateral mediastinal LN metastasis; M1: distant metastasis or malignant effusion).
<b>s</b>. Supplemental Table 2 of @TLB+16, filtered with adjusted _p_-value < 0.05 and fold change ≥ 1.5 or ≤ 2/3.
<b>t</b>. Gene names from Supplementary Table S1 of @FGW+17, filtered with _p_-value < 0.05.
<b>u</b>. Table 1 of @LZW+17.
<b>v</b>. Supplementary Table S2 of @SFS+17 (sheet LF: label-free quantification).
<b>w</b>. Supplementary Table 1 of @WLC+17.
<b>x</b>. <b>y</b>. <b>z</b>. <b>A</b>. <b>B</b>. <b>C</b>. Gene names from Tables S2--S6 of @YCC+17 filtered to keep proteins with fold change ≥1.5 or ≤2/3 (Oncogene: oncogene-coded proteins; TSG: tumor suppressor gene-coded proteins; Glycoprotein: glycoproteomics data).
<b>D</b>. Gene names from Supplementary Table S2 of @NLW+18.

## References
