---
title: "More about metrics"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{More about metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: canprot.bib
link-citations: yes
csl: elementa.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
oldopt <- options(width = 72)
```

```{r HTML, include=FALSE}
Zc <- "<i>Z</i><sub>C</sub>"
nC <- "<i>n</i><sub>C</sub>"
nH2O <- "<i>n</i><sub>H<sub>2</sub>O</sub>"
H2O <- "H<sub>2</sub>O"
O2 <- "O<sub>2</sub>"
```

The **canprot** package has functions to calculate chemical metrics of proteins from their amino acid composition.
This vignette was compiled on `r Sys.Date()` with **canprot** version `r packageDescription("canprot")$Version`.

Previous vignette: [Introduction to canprot](introduction.html)

## More details on chemical metrics

Carbon oxidation state (`r Zc`) is calculated from elemental ratios, but stoichiometric hydration state (`r nH2O`) depends on a specific choice of thermodynamic components (or basis species).
By default, `r nH2O` is calculated from a theoretical reaction to form proteins from this set of basis species, abbreviated as **QEC**:

* glutamine (Q)
* glutamic acid (E)
* cysteine (C)
* `r H2O`
* `r O2`

The rationale for this choice is described in @DYT20 and @Dic22.

To see how it works, consider the formation reaction of alanylglycine, which can be written using functions in the **CHNOSZ** package:
```{r AG_reaction, message = FALSE}
CHNOSZ::basis("QEC")
CHNOSZ::subcrt("alanylglycine", 1)$reaction
```

As it turns out, alanylglycine has the same elemental formula as glutamine, which is one of the basis species, so there are no other basis species in the reaction.
That includes `r H2O`, so we can deduce that `r nH2O` is zero.

Let's do the calculation with the `nH2O()` function to see this result.
We have to specify `terminal_H2O = 1` in order to account for the terminal -H and -OH groups on alanlglycine.
As described below, the default for this setting is 0 because, most of the time, we want to measure the per-amino-acid contributions for proteins.
```{r AG_metrics}
AG <- data.frame(Ala = 1, Gly = 1)
nH2O(AG, terminal_H2O = 1)
```

Now let's try an actual protein, chicken egg-white lysozome, which has the name LYSC\_CHICK in UniProt with accession number [P00698](https://www.uniprot.org/uniprotkb/P00698).
The amino acid compositions of this and selected other proteins are available in the **CHNOSZ** package.
This gets the amino acid composition and prints the protein length and `r Zc` and `r nH2O`:
```{r LYSC_CHICK_metrics}
AA <- CHNOSZ::pinfo(CHNOSZ::pinfo("LYSC_CHICK"))
plength(AA)
Zc(AA)
nH2O(AA)
```

*By the way, lysozyme (a secreted protein) is highly oxidized compared to cytoplasmic proteins, most of which have negative `r Zc`.*

To see where the value of `r nH2O` comes from, we can write the formation reaction of LYSC\_CHICK from the **QEC** basis species.
Recall that we set the basis species above with `CHNOSZ::basis("QEC")`; that setting is used here to balance the reaction.
```{r LYSC_reaction, message = FALSE, echo = 1:2}
(reaction <- CHNOSZ::subcrt("LYSC_CHICK", 1)$reaction) # print the reaction
(H2Ocoeff <- with(reaction, coeff[name == "water"])) # print the coefficient on H2O
stopifnot( -(H2Ocoeff + 1) / plength(AA) == nH2O(AA))
```

That says `r H2Ocoeff`, but the value for `r nH2O` above was `r nH2O(AA)`.
What happened?
Let's step through the logic:

* The reaction shows that `r H2Ocoeff` water molecules are *released* (the coefficient is positive) in the theoretical formation of one molecule of the protein.
* The protein has terminal -H and -OH groups. This means that shorter proteins are "wetter", and we don't want that complication. So, we take one `r H2O` away from the protein and add it to the number of water molecules released, giving `r H2Ocoeff` + 1 = `r H2Ocoeff + 1` `r H2O`. _This step was not performed by @DYT20 but is the default in **canprot** since version 2.0.0._
* We use the opposite of this value because we are counting how many `r H2O` units go **into** forming the protein.
* Finally, we divide by the length of the protein to get the stoichiometric hydration state normalized per residue: `r -(H2Ocoeff + 1)` / `r plength(AA)` = `r nH2O(AA)`

A general observation: This is a theoretical reaction in terms of thermodynamic components, so we are not dealing with biochemical mechanisms here.
That's one reason for calling this approach *geochemical biology*.

## Implementation details

To save time, `nH2O()` does not calculate the formation reaction for each protein but instead uses precomputed values of `r nH2O` for each amino acid.
The two methods give equivalent results, as described in @DYT20.

Similarly, `Zc()` uses precomputed values of `r Zc` and `r nC` (number of carbon atoms) for each amino acid.
**NOTE**: Calculating `r Zc` of proteins from amino acid frequencies (i.e. abundances or counts in a protein sequence) requires weighting the amino-acid `r Zc` by the number of carbon atoms in each amino acid, in addition to weighting by amino acid frequency.
Using the unweighted mean of `r Zc` of amino acids leads to artificially higher values for proteins.

## GRAVY, pI, and molecular weight

There are also functions for calculating the grand average of hydropathicity (GRAVY) and isoelectric point (pI) of proteins.
Below, values for representative proteins are compared with results from the ProtParam tool [@GHG+05] in UniProt.

We first look up a few proteins in CHNOSZ's list of proteins, then get the amino acid compositions.
```{r CHNOSZ_proteins}
iprotein <- CHNOSZ::pinfo(c("LYSC_CHICK", "RNAS1_BOVIN", "AMYA_PYRFU", "CSG_HALJP"))
AAcomp <- CHNOSZ::pinfo(iprotein)
```

Calculate GRAVY and compare it to reference values from <https://web.expasy.org/protparam/>.
```{r GRAVY}
G_calc <- GRAVY(AAcomp)
# https://web.expasy.org/cgi-bin/protparam/protparam1?P00698@19-147@
# https://web.expasy.org/cgi-bin/protparam/protparam1?P61823@27-150@
# https://web.expasy.org/cgi-bin/protparam/protparam1?P49067@2-649@
G_ref <- c(-0.472, -0.663, -0.325)
stopifnot(all.equal(round(G_calc[1:3], 3), G_ref, check.attributes = FALSE))
```

Calculate pI and compare it to reference values.
```{r pI}
pI_calc <- pI(AAcomp)
# Reference values calculated with ProtParam
# LYSC_CHICK: residues 19-147 (sequence v1)
# RNAS1_BOVIN: residues 27-150 (sequence v1)
# AMYA_PYRFU: residues 2-649 (sequence v2)
# CSG_HALJP: residues 35-862 (sequence v1)
pI_ref <- c(9.32, 8.64, 5.46, 3.37)
stopifnot(all.equal(as.numeric(pI_calc), pI_ref))
```

Calculate molecular weight (`MW`) and compare it to reference values
```{r MW}
# Per-residue molecular weight multiplied by number of residues
MWcalc <- MW(AAcomp) * plength(AAcomp)
# Add terminal groups
MWcalc <- MWcalc + 18.01528
# Reference values for molecular weights of proteins
MWref <- c(14313.14, 13690.29, 76178.25)
stopifnot(all.equal(round(MWcalc[1:3], 2), MWref, check.attributes = FALSE))
```

```{r reset, include=FALSE}
options(oldopt)
```

## References
