---
title: "Prostate Cancer"
output:
  html_vignette:
    mathjax: null
vignette: >
  %\VignetteIndexEntry{Prostate Cancer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: cpdat.bib
csl: peerj.csl
---

```{r setup, include=FALSE}
library(canprot)
library(CHNOSZ)
library(knitr)
## use pngquant to reduce size of PNG images
knit_hooks$set(pngquant = hook_pngquant)
pngquant <- "--speed=1 --quality=0-25"
# in case pngquant isn't available (R-Forge?)
if (!nzchar(Sys.which("pngquant"))) pngquant <- NULL 
```

This vignette shows compositional metrics and phylostrata for proteins that are differentially expressed in prostate cancer compared to normal tissue.
Abbreviations:

  * <i>Z</i><sub>C</sub> &ndash; carbon oxidation state; <i>n</i><sub>H<sub>2</sub>O</sub> &ndash; stoichiometric hydration state; <i>n</i><sub>O<sub>2</sub></sub> &ndash; stoichiometric oxidation state.
  * <i>n</i><sub>AA</sub> &ndash; protein length; PS &ndash; phylostrata.

Stoichiometric values are derived from basis species using the rQEC method (Dick et al., 2020) or from amino acid biosynthetic reactions.
References for gene ages: [Trigos et al. (2017)](https://doi.org/10.1073/pnas.1617743114) (TPPG17); [Liebeskind et al. (2016)](https://doi.org/10.1093/gbe/evw113) (LMM16).

```{r options, echo=FALSE}
options(width = 90)
```

<style type="text/css">
body {
  max-width: 800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r datasets}
datasets <- pdat_prostate()
```

```{r comptab, results="hide", message=FALSE, echo = FALSE}
pdat1 <- lapply_canprot(datasets, pdat_prostate)
comptab1 <- lapply(pdat1, get_comptab)
pdat2 <- lapply(pdat1, pdat_recomp, "biosynth")
comptab2 <- lapply(pdat2, get_comptab, "nO2")
comptab3 <- lapply(pdat1, get_comptab, "nAA", "PS")
comptab4 <- lapply(pdat1, get_comptab, "nAA", "PS", PS_source = "LMM16")
```

Dashed contour lines in the plots outline the 50% credible region for highest probability density.
Triangle symbols indicate transcriptomic datasets; all others are protein expression.

```{r diffplot, fig.width=8, fig.height=8, fig.align = "center", echo = FALSE, pngquant = pngquant}
par(mfrow = c(2, 2), mar = c(4, 4, 2, 2), mgp = c(2.5, 1, 0))
pch <- ifelse(grepl("transcript", datasets), 2, 1)
diffplot(comptab1, pch = pch)
title(quote("rQEC"~italic(n)[H[2]*O]), font.main = 1)
diffplot(comptab2, c("nO2", "nH2O"), pch = pch)
title(quote("Biosynthetic"~italic(n)[H[2]*O]~"and"~italic(n)[O[2]]), font.main = 1)
diffplot(comptab3, c("nAA", "PS"), pch = pch)
title("Trigos et al. (2017) ages", font.main = 1)
diffplot(comptab4, c("nAA", "PS"), pch = pch)
title("Liebeskind et al. (2016) ages", font.main = 1)
```

In the table, values of &Delta;<i>Z</i><sub>C</sub>, &Delta;<i>n</i><sub>H<sub>2</sub>O</sub> and &Delta;<i>n</i><sub>O<sub>2</sub></sub> are multiplied by 1000, values of &Delta;PS are multiplied by 100, and negative values are shown in bold.
Abbreviations:

  * <i>n</i><sub>down</sub> &ndash; number of down-expressed proteins; <i>n</i><sub>up</sub> &ndash; number of up-expressed proteins.
  * PCa &ndash; prostate cancer; BPH &ndash; benign prostatic hyperplasia.
  * FFPE &ndash; formalin-fixed paraffin-embedded; OCT &ndash; optimal cutting temperature compound-embedded; TMA &ndash; tissue microarray.

```{r xsummary, results="asis", echo = FALSE}
library(xtable)
out <- xsummary2(comptab1, comptab2, comptab3, comptab4)
# round values and include dataset tags
tags <- sapply(sapply(strsplit(datasets, "="), "[", -1), paste, collapse = ";")
out <- cbind(out[, 1:2], tags = tags, out[, 3:11])
out[, 6:12] <- round(out[, 6:12], 4)
write.csv(out, "prostate.csv", row.names = FALSE, quote = 2)
```

## Data Sources
Gene names or other identifiers were converted to UniProt accession numbers using the <a href="https://www.uniprot.org/mapping/">UniProt mapping tool</a>.

<b>a</b>. Table 2 of @GTR+08.
<b>b</b>. Supplementary Tables 3 and 4 of @KPB+10.
<b>c</b>. <b>d</b>. Table S1 (genes) and S2 (proteins) of @HZH+12.
<b>e</b>. Table S1 of @JHZ+13.
<b>f</b>. Supplemental Table S7 of @LCS+14.
<b>g</b>. Supplementary Table 1 @CZL+16 (proteins recorded with only Down-regulation or Up-regulation).
<b>h</b>. Supplementary Table S3 of @IWT+16, filtered to include proteins listed with FDR <0.1.
<b>i</b>. <b>j</b>. Extracted from Table S4 @GLZ+18. Values were quantile normalized (R function preprocessCore::normalize.quantiles) ([Bolstad et al., 2003](https://doi.org/10.1093/bioinformatics/19.2.185)), then ratios were calculated between the median values for each cancer type (acinar and ductal) and corresponding normal tissue; ratios > 1.5 or < 2/3 were used to identify differentially expressed proteins.
<b>k</b>. <b>l</b>. Extracted from Supplementary Data 1 of @LAJ+18 (sheet "Area-proteins") by applying quantile normalization to peak areas then calculating median values across all runs and samples for each of BPH , PC (primary prostate cancer), and CRPC (castration resistant prostate cancer). A cutoff of 2-fold in ratios of medians (PC / BPH or CRPC / BPH) was used to identify differentially expressed proteins.
<b>m</b>. Table 2 of @MAN+18.
<b>n</b>. <b>o</b>. <b>p</b>. <b>q</b>.  <b>r</b>. Supporting Information Table S3 of @KRN+19 (G1&ndash;G5: PCa grades).
<b>s</b>. <b>t</b>. Tables S4d (GS = 6) and S4f (GS ≥ 8 or GS = 6) of @MMF+19 (GS: Gleason Score).
<b>u</b>. Supplementary Table S3a of @TOT+19.
<b>v</b>. <b>w</b>. Table S5 of @ZYW+19 (LG: low-grade PCa; HG: high-grade PCa).

## References
